---
title: "Gender SPI Correlation Analysis"
author: "Brian Stacy"
date: "7/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	fig.height=5,
	fig.width=7
)
options(rgl.useNULL = TRUE)
library(tidyverse)
library(haven)
library(lattice)
library(here)
library(ggpmisc)
library(ggthemes)
library(ggtext)
library(estimatr)
library(flextable)

#set directories
dir <- here()

raw_dir <- paste(dir, '01_raw_data', sep="/")
output_dir <- paste(dir, '03_output_data', sep="/")
```

```{r data}
window <- 5
pop_filt <- 200000

 #add in population
pop_df <- wbstats::wb_data(country="all",
             indicator='SP.POP.TOTL',
             start_date=2004,
             end_date=2021) %>%
  mutate(date=as.numeric(date)) %>%
  mutate(population=SP.POP.TOTL) %>%
  select(iso3c, date, population) 


#get a list of countries that are below the population limit defined at to of script
pop_exlude_cntry <- pop_df %>% filter(date==2021) %>% filter(population <= pop_filt) %>% select(iso3c) %>% pluck(1)
pop_exlude_cntry_names <- pop_df %>% left_join(wbstats::wb_countries()) %>% filter(date==2021) %>% filter(population <= pop_filt) %>% select(country) %>% pluck(1)


#get overall gender data availability measure
avail_SPI_df <- read_csv(paste0(output_dir, '/SPI_Gender_UNSD_data_',window,'yr.csv')) %>%   
  filter(!(iso3c %in% pop_exlude_cntry)) %>%
  
  mutate(availability=100*ind_quality) %>% #convert ind_quality measure which is 0-1 to percentage available by multiplying by 100
  filter(date==2019 & goal=="Combined") #keep just most recent year

#read in SPI data from WDI
SPI_df <- wbstats::wb_data(indicator=c('IQ.SPI.OVRL','IQ.SPI.PIL1', 'IQ.SPI.PIL2','IQ.SPI.PIL3','IQ.SPI.PIL4','IQ.SPI.PIL5',
                                       'NY.GDP.PCAP.PP.CD', 'NY.GDP.MKTP.PP.CD','SP.POP.TOTL', 'SG.LAW.INDX',
                                       'HD.HCI.OVRL','GE.EST'),
                           start_date = 2019) 




#combine data
combined_df <- avail_SPI_df %>%
  left_join(SPI_df) %>%
  filter(!is.na(IQ.SPI.OVRL))

#add equations to plots
eq_plot_txt <- function(data) {
  eq <- lm_robust(outcome ~ availability, data=data, se_type='HC2')
  coef <- coef(eq)
  std_err <- sqrt(diag(vcov(eq)))
  r_2<- summary(eq)$r.squared
  sprintf("y = %.3f + %.3f x, R<sup>2</sup> = %.3f <br>    (%.3f)   (%.3f)", coef[1], coef[2], r_2[1], std_err[1], std_err[2])

}

#add equations to plots
eq_plot_txt_reverse <- function(data, inp, var) {
  eq <- lm_robust(inp ~ var, data=data, se_type='HC2')
  coef <- coef(eq)
  std_err <- sqrt(diag(vcov(eq)))
  r_2<- summary(eq)$r.squared
  sprintf("y = %.3f + %.3f x, R<sup>2</sup> = %.3f <br>    (%.3f)   (%.3f)", coef[1], coef[2], r_2[1], std_err[1], std_err[2])

}

```


# Introduction

```{r}



gend_gdp_plt <-combined_df %>%
  ggplot(aes(x=NY.GDP.PCAP.PP.CD, y=availability)) +
  geom_text(aes(label=iso3c)) +
  geom_smooth(method='lm') +
  theme_bw() + 
  scale_x_log10(labels=scales::comma) +
  xlab('GDP per capita, PPP (current international $), logged') +
  ylab('Availability of Gender SDG Indicators (%)') +
  geom_richtext(
    aes(y = 87, x = 10000,label = eq_plot_txt_reverse(combined_df,availability, log10(NY.GDP.PCAP.PP.CD)), hjust=0.2)
  ) +  
  theme(legend.position = 'bottom') +
  expand_limits(y=c(0,100)) +
  labs(
    title='Plot of GDP per capita and Availability of Gender SDG Indicators',
    caption=str_wrap('Source: GDP per capita from the World Development Indicatrs (NY.GDP.PCAP.PP.CD).  The availability of Gender SDG Indicators based on authors calculations based on the UN Global SDG Database.',100)
  )

gend_gdp_plt


```

```{r}



gend_gdp_plt <-combined_df %>%
  ggplot(aes(x=NY.GDP.MKTP.PP.CD, y=availability)) +
  geom_text(aes(label=iso3c)) +
  geom_smooth(method='lm') +
  theme_bw() + 
  scale_x_log10(labels=scales::comma) +
  xlab('GDP, PPP (current international $), logged') +
  ylab('Availability of Gender SDG Indicators (%)') +
  geom_richtext(
    aes(y = 87, x = 1000000000,label = eq_plot_txt_reverse(combined_df,availability, log10(NY.GDP.MKTP.PP.CD)), hjust=0.2)
  ) +  
  theme(legend.position = 'bottom') +
  expand_limits(y=c(0,100)) +
  labs(
    title='Plot of GDP and Availability of Gender SDG Indicators',
    caption=str_wrap('Source: GDP from the World Development Indicatrs (NY.GDP.MKTP.PP.CD).  The availability of Gender SDG Indicators based on authors calculations based on the UN Global SDG Database.',100)
  )

gend_gdp_plt


```

```{r}
spi_gdp_plt <-combined_df %>%

  ggplot(aes(x=NY.GDP.PCAP.PP.CD, y=IQ.SPI.OVRL)) +
  geom_text(aes(label=iso3c)) +
  geom_smooth(method='lm') +
  theme_bw() + 
  scale_x_log10(labels=scales::comma) +
  xlab('GDP per capita, PPP (current international $), logged') +
  ylab('SPI Overall Score') +
  geom_richtext(
    aes(y = 10, x = 10000,label = eq_plot_txt_reverse(combined_df,IQ.SPI.OVRL, log10(NY.GDP.PCAP.PP.CD)), hjust=0.2)
  ) +  
  theme(legend.position = 'bottom') +
  expand_limits(y=c(0,100)) +
  labs(
    title='Plot of GDP per capita and SPI Overall Score',
    caption=str_wrap('Source: GDP per capita from the World Development Indicatrs (NY.GDP.PCAP.PP.CD).  SPI Overall Score from the World Development Indicators (IQ.SPI.OVRL).',100)
  )

spi_gdp_plt


```


```{r}
spi_gdp_plt <-combined_df %>%

  ggplot(aes(x=NY.GDP.MKTP.PP.CD, y=IQ.SPI.OVRL)) +
  geom_text(aes(label=iso3c)) +
  geom_smooth(method='lm') +
  theme_bw() + 
  scale_x_log10(labels=scales::comma) +
  xlab('GDP, PPP (current international $), logged') +
  ylab('SPI Overall Score') +
  geom_richtext(
    aes(y = 10, x = 10000000000,label = eq_plot_txt_reverse(combined_df,IQ.SPI.OVRL, log10(NY.GDP.MKTP.PP.CD)), hjust=0.2)
  ) +  
  theme(legend.position = 'bottom') +
  expand_limits(y=c(0,100)) +
  labs(
    title='Plot of GDP and SPI Overall Score',
    caption=str_wrap('Source: GDP from the World Development Indicatrs (NY.GDP.MKTP.PP.CD).  SPI Overall Score from the World Development Indicators (IQ.SPI.OVRL).',100)
  )

spi_gdp_plt


```

```{r}
#plot SPI scores against availability
gend_spi_plt <-combined_df %>%
  ggplot(aes(x=IQ.SPI.OVRL, y=availability)) +
  geom_text(aes(label=iso3c)) +
  geom_smooth(method='lm') +
  theme_bw() +
  xlab('SPI Overall Score') +
  ylab('Availability of Gender SDG Indicators (%)') +
  geom_richtext(
    aes(x = 75, y = 5,label = eq_plot_txt_reverse(combined_df,availability, IQ.SPI.OVRL), hjust=0.2)
  ) +  
  theme(legend.position = 'bottom') +
  expand_limits(x=c(0,100),
                y=c(0,60)) +
  labs(
    title='Plot of SPI overall score and Availability of Gender SDG Indicators',
    caption=str_wrap('Source: SPI Overall Score from the World Development Indicators (IQ.SPI.OVRL).  The availability of Gender SDG Indicators based on authors calculations based on the UN Global SDG Database.',100)
  )

gend_spi_plt
```



# Correlations

Can you send the correlation between GDP PC and each of the 5 subcomponents of the SPI?  And then also send the correlation between the G SPI and each of the 5 subcomponents of the SPI? Along with the 3 main correlations (GDP pc & SPI, GPD pc & G SPI, and SPI & G SPI). So 13 correlations total.

Source: GDP per capita, PPP (current international $) from the World Development Indicatrs (NY.GDP.PCAP.PP.CD).  SPI Overall Score from the World Development Indicators.  GDP is logged.  *** is significant at 0.001 level. ** is significant at 0.01 level. * is significant at 0.05 level. + is significant at 0.01 level. 

```{r}

cor_fun  <- function(var1, var2) {
  
  data <- combined_df %>%
    rename(
      x=!! var1,
      y=!! var2
    )
  
  r <- Hmisc::rcorr(log(data$x),data$y)
  
  pval <- r$P[1,2]
  
  if (pval<0.001) {
    stars<-"***"
  } 
  else if (pval<0.01) {
    stars<-"**"
  } 
  else if  (pval<0.05) {
    stars<-"*"
  } 
  else if  (pval<0.1) {
    stars<-"+"
  }   
  else {
    stars <- ""
  }
  
  paste0(round(r$r[1,2],2),stars)
  
}

cor_fun_log  <- function(var1, var2) {
  
  data <- combined_df %>%
    rename(
      x=!! var1,
      y=!! var2
    )
  
  r <- Hmisc::rcorr(log(data$x),data$y)
  
  pval <- r$P[1,2]
  
  if (pval<0.001) {
    stars<-"***"
  } 
  else if (pval<0.01) {
    stars<-"**"
  } 
  else if  (pval<0.05) {
    stars<-"*"
  } 
  else if  (pval<0.1) {
    stars<-"+"
  }     
  else {
    stars <- ""
  }
  
  paste0(round(r$r[1,2],2),stars)
  
}
```

## GDP per capita Correlations

 Log GDP per capita & Gender SDG Availability: `r cor_fun_log('NY.GDP.PCAP.PP.CD', 'availability')`.

 Log GDP per capita & SPI Overall Score: `r cor_fun_log('NY.GDP.PCAP.PP.CD', 'IQ.SPI.OVRL')`.

 Log GDP per capita & SPI Pillar 1 (Data use) Score: `r cor_fun_log('NY.GDP.PCAP.PP.CD', 'IQ.SPI.PIL1')`.

 Log GDP per capita & SPI Pillar 2 (Data services) Score: `r cor_fun_log('NY.GDP.PCAP.PP.CD', 'IQ.SPI.PIL2')`.

 Log GDP per capita & SPI Pillar 3 (Data products [Overall SDG Indicator availability]) Score: `r cor_fun_log('NY.GDP.PCAP.PP.CD', 'IQ.SPI.PIL3')`.

 Log GDP per capita & SPI Pillar 4 (Data sources) Score: `r cor_fun_log('NY.GDP.PCAP.PP.CD', 'IQ.SPI.PIL4')`.

 Log GDP per capita & SPI Pillar 5 (Data infrastructure) Score: `r cor_fun_log('NY.GDP.PCAP.PP.CD', 'IQ.SPI.PIL5')`.

## SPI Correlations


 Gender SDG Availability & SPI Overall Score: `r cor_fun('availability', 'IQ.SPI.OVRL')`.

 Gender SDG Availability & SPI Pillar 1 (Data use) Score: `r cor_fun('availability', 'IQ.SPI.PIL1')`.

 Gender SDG Availability & SPI Pillar 2 (Data services) Score: `r cor_fun('availability', 'IQ.SPI.PIL2')`.

 Gender SDG Availability & SPI Pillar 3 (Data products [Overall SDG Indicator availability]) Score: `r cor_fun('availability', 'IQ.SPI.PIL3')`.

 Gender SDG Availability & SPI Pillar 4 (Data sources) Score: `r cor_fun('availability', 'IQ.SPI.PIL4')`.

 Gender SDG Availability & SPI Pillar 5 (Data infrastructure) Score: `r cor_fun('availability', 'IQ.SPI.PIL5')`.
 
 
# Regression Analysis

```{r}
#examine whether residuals of regression of SPI overall score on GDP per capita are strongly correlated with Gender SDG availability

mod <- lm(IQ.SPI.OVRL ~ NY.GDP.PCAP.PP.CD, combined_df)

resid <- resid(mod)
pred <- predict(mod)
avail <- combined_df %>% filter(!is.na(NY.GDP.PCAP.PP.CD)) %>% pull(availability)

cor(resid, avail)
cor(pred, avail)


summary(lm(availability ~ log10(NY.GDP.PCAP.PP.CD) + IQ.SPI.OVRL, combined_df))
summary(lm(availability ~  IQ.SPI.OVRL, combined_df))
summary(lm(availability ~  log(SP.POP.TOTL), combined_df))


```
 
 # SDG 5
 
 For the 14 SDG 5 indicators, can you send me the distribution of countries with only 1, only 2, 3, 4, … up to all 14? 
So the %s will add up to 100. And remind me of the N (number of countries)

```{r sdggendf}
#read in table about gender sdgs
sdg_tab <- read_csv(file = paste0(raw_dir, '/misc/Gender_SDGs.csv')) 

sdg_tab_merge <- sdg_tab %>%
  #filter(!is.na(Code)) %>%
  transmute(ccode=ccode,
            code1=Code,
            code=SDG,
            Gender_Breakdown=Gender_Breakdown)

#get a list of all SDG indicators
indicators_url <- paste0(raw_dir, '/misc/un_indicators.json')

indicators_list <- sdg_tab %>%
  filter(!is.na(Code))

indicators_list <- indicators_list$Code


#bring in the list of indicators
list_df <- jsonlite::fromJSON(indicators_url, flatten = TRUE) %>%
  as_tibble() %>%
  unnest(keep_empty = TRUE)

list_df <- sdg_tab_merge %>%
  left_join(list_df) %>%
  #filter(tier<=2) %>% # keep just the tier 1 indicators %>%
  #filter(code1 %in% indicators_list) 
  filter(!is.na(ccode)) 

# get list of tier 1 indicaors
tier1_list <- list_df %>%
  filter(tier==1)

tier1_list <- tier1_list$code1

# get list of tier 2 indicaors
tier2_list <- list_df %>%
  filter(tier==2)

tier2_list <- tier2_list$code1

#get goal
sdg_tab$Goal <- as.character(map(strsplit(sdg_tab$SDG, split = "\\."), 1))

gend_df_len <- nrow(list_df)

country_metadata <- wbstats::wb_countries() %>%
  filter(region!="Aggregates")

```

```{r}
date_start=2016
date_end=2020

sdg_5_indicators_df <- read_csv(paste(output_dir, '/un_sdg_combined_df.csv',sep="")) %>%
  filter(goal=="5")

  
sdg_5_indicators_df <- sdg_5_indicators_df %>%
    filter(between(date,date_start,date_end) ) %>%
    filter(!is.na(ind_quality)) %>%
    group_by(iso3c, ccode) %>%
    summarise(ind_quality=max(ind_quality, na.rm=T),
              ind_value=mean(ind_value, na.rm=T),
              ind_metadata=first(ind_metadata)) %>% #check if any values (even sub-indicators) for indicator
      mutate(ind_quality=if_else(is.na(ind_quality),0,ind_quality)) %>% # if the indicator is missing for a year in the database, set availability to 0.
      left_join(list_df %>% group_by(ccode,code, goal) %>% summarise(num=n())) %>%
    left_join(country_metadata) %>%
    filter(!is.na(region)) %>%
    select(iso3c, country,code, ccode, region, income_level,goal, ind_quality) %>%
    mutate(date=date_end
           ) 

sdg_5_tally <- sdg_5_indicators_df %>%
  filter(date==2020) %>%
  group_by(country, iso3c, region ) %>%
  summarise(n=n(),
            SDGs=sum(ind_quality>0),
            SDG_share=100*SDGs/14)
  

```

```{r}
ggplot(sdg_5_tally, aes(x=SDGs)) +
  geom_bar() +
  theme_bw() +
  ggtitle('Distribution of SDG 5 Availability from 2016-2020 Across Countries')


share <- sdg_5_tally %>%
  group_by(SDGs) %>%
  summarise(count=n(),
            share=100*count/215) 


ggplot(share, aes(x=SDGs, y=cumsum(share))) + 
  geom_line() + 
  ylab('Percent') +
  xlab('Number of SDG 5') +
  geom_point() +
  theme_bw() +
  ggtitle('Cumulative Share of SDG 5 Availability from 2016-2020 Across Countries')  


```


```{r}
share %>%
  transmute(
    `Number of SDG 5`=SDGs,
    `Number of Countries`=count,
    `Share`=share
  ) %>%
  flextable() %>%
  add_header_lines('Distribution of SDG 5 Availability from 2016-2020 Across Countries') %>%
  autofit()

```

